Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-03-21 paper/ms.tex
more words aboutpostivie definiteness

To scale GP models to larger datasets, \citet{Rybicki:1995} presented a method
of computing the first term in \eq{gp-likelihood} in $\mathcal{O}(N)$
operations when the covariance function is given by
\begin{eqnarray}\eqlabel{kernel-simple}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} + a\,\exp(-c\,\tau_{nm})
\end{eqnarray}
where $\{{\sigma_n}^2\}_{n=1}^N$ are the measurement uncertainties,
$\delta_{nm}$ is the Kronecker delta, and $\bvec{\alpha} = (a,\,c)$.
The intuition behind this method is that, for this choice of $k_\bvec{\alpha}$,
the inverse of $K_\bvec{\alpha}$ is tridiagonal and can be computed
with a small number of operations for each data point.
Subsequently, \citet{Ambikasaran:2015} generalized this method to arbitrary
mixtures of exponentials
\begin{eqnarray}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} +
    \sum_{j=1}^J a_j\,\exp(-c_j\,\tau_{nm})\quad.
\end{eqnarray}
In this case, the inverse will be dense but \eq{gp-likelihood} can still be
evaluated in $\mathcal{O}(N)$ operations where $J$ is the number of components
in the mixture and $N$ is still the number of data points.

@@

In the case of the full \celerite\ covariance function
(\eqalt{celerite-kernel}), we introduce the following auxiliary variables in
analogy to the $u_n$ and $g_n$ that we introduced in the previous section
\begin{eqnarray}\eqlabel{full-system-first}
\phi_{n,j} &=& e^{-c_j\,\tau_{n+1,n}}\,\cos\left(d_j\,\tau_{n+1,n}\right)\\
\psi_{n,j} &=& -e^{-c_j\,\tau_{n+1,n}}\,\sin\left(d_j\,\tau_{n+1,n}\right)\\
g_{n,j} &=& \phi_{n,j}\,g_{n-1,j} + \phi_{n,j}\,z_n + \psi_{n,j}\,h_{n-1,j}\\
h_{n,j} &=& \phi_{n,j}\,h_{n-1,j} - \psi_{n,j}\,z_n - \psi_{n,j}\,g_{n-1,j}\\
u_{n,j} &=& \phi_{n+1,j}\,u_{n+1,j} + a_j\,z_{n+1} + \psi_{n+1,j}\,v_{n+1,j}\\
v_{n,j} &=& \phi_{n+1,j}\,v_{n+1,j} - b_j\,z_{n+1} - \psi_{n+1,j}\,u_{n+1,j}
\end{eqnarray}
with the boundary conditions
\begin{eqnarray}
    g_{0,j} = 0 \quad, \quad
    h_{0,j} = 0 \quad, \quad
    u_{N,j} = 0 \quad, \quad\mathrm{and}\quad
    v_{N,j} = 0
\end{eqnarray}
for all $j$.
Using these variables and some algebra, we find that the following expression
\begin{eqnarray}\eqlabel{full-system-last}
\sum_{j=1}^J \left[a_j\,g_{n,j}+b_j\,h_{n,j}\right]
+ \left[{\sigma_n}^2+\sum_{j=1}^J a_j\right]
+ \sum_{j=1}^J \left[\phi_{n,j}\,u_{n,j}+\psi_{n,j}\,v_{n,j}\right]
    &=& r_{\bvec{\theta},n}
\end{eqnarray}
is equivalent to the target matrix equation
\begin{eqnarray}
K\,\bvec{z} &=& \bvec{r}_\bvec{\theta}
\end{eqnarray}
if $r_{\bvec{\theta},n}$ is the $n$-th element of the residual vector
$\bvec{r}_\bvec{\theta}$ defined in \sect{gps}.
\eq{full-system-first} through \eq{full-system-last} define a banded matrix
equation in the ``extended'' space and, as before, this can be used to solve
for $K^{-1}\,\bvec{r}_\bvec{\theta}$ and $\det K$ in $\mathcal{O}(N)$
operations.
\Figure{matrix} shows a pictorial representation of the sparsity pattern of
the extended matrix $K_\mathrm{ext}$.
Given that definition of $K_\mathrm{ext}$, the corresponding extended vectors
$\bvec{z}_\mathrm{ext}$ and $\bvec{r}_\mathrm{ext}$ are defined schematically
as
\begin{eqnarray}
{\bvec{z}_\mathrm{ext}} ^\T =
\left(\begin{array}{cccccccccc}
    z_1 & u_{1,j} & v_{1,j} & g_{1,j} & h_{1,j} & z_2 & u_{2,j} & \cdots &
    h_{N-1,j} & z_N
\end{array}\right)
\end{eqnarray}
and
\begin{eqnarray}
{\bvec{r}_\mathrm{ext}} ^\T =
\left(\begin{array}{cccccccccc}
    r_{\bvec{\theta},1} & 0 & 0 & 0 & 0 & r_{\bvec{\theta},2} &
    0 & \cdots & 0 & r_{\bvec{\theta},N}
\end{array}\right) \quad.
\end{eqnarray}
After constructing the extended matrix (using a compact storage format), the
extended matrix can be factorized using a LU-decomposition\footnote{Even
though $K_\mathrm{ext}$ is symmetric, it is not positive definite so a
Cholesky solver cannot be used for increased efficiency.} routine optimized
for band or sparse matrices.
This decomposition can then be used to compute the determinant of $K$, solve
$K^{-1}\,\bvec{r}_\bvec{\theta}$, and subsequently calculate the marginalized
likelihood in \eq{gp-likelihood}.

@@

The extended system defined in the previous section is sparse with typically fewer than
a few percent non-zero entries and band structure.
In this \sectionname, we empirically investigate the performance and scaling
of three different algorithms for solving this extended system:

@@

We now turn to the application of the \celerite\ algorithm to astrophysical
variability.  Many astronomical objects have a characteristic timescale
which is determined by the physical structure.  Variability
is excited by noisy physical processes, which grows most strongly at
the characteristic timescale, but is also damped due to dissipation 
in the system.  As a concrete example, stellar oscillations are strong at 
resonant frequencies determined by the internal stellar structure,
which are both excited and damped by convective turbulence.

Thus, a special case of the \celerite\ model of great physical interest is a
stochastically-driven damped simple harmonic oscillator.
The differential equation for this system is
\begin{equation}
    \left[\frac{\dd^2}{\dd t^2} + \frac{\omega_0}{Q}\,\frac{\dd}{\dd t}
    + \omega_0^2\right]\, y(t) = \epsilon(t)
\end{equation}
where $\omega_0$ is the frequency of the undamped oscillator, $Q$ is the
quality factor of the oscillator, and $\epsilon(t)$ is a stochastic driving
force.
If $\epsilon(t)$ is white noise, the PSD of this process is given by
\citep[for example][]{Anderson:1990}
\begin{equation}\eqlabel{sho-psd}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0\,\omega_0^4}
    {(\omega^2-\omega_0^2)^2 + \omega_0^2\omega^2/Q^2}
\end{equation}
where $S_0$ is proportional to the power at $\omega = \omega_0$, $S(\omega_0)
= \sqrt{2/\pi}\,S_0\,Q^2$.
The power spectrum in \eq{sho-psd} matches \eq{celerite-psd} if
\begin{eqnarray}\eqlabel{sho-complex}
a_j &=& S_0\,\omega_0\,Q \\
b_j &=& \frac{S_0\,\omega_0\,Q}{\sqrt{4\,Q^2-1}} \\
c_j &=& \frac{\omega_0}{2\,Q}\\
d_j &=& \frac{\omega_0}{2\,Q} \sqrt{4\,Q^2-1} \quad,
\end{eqnarray}
for $Q \ge \frac{1}{2}$.
For $0 < Q \le \frac{1}{2}$, \eq{sho-psd} can be captured by a pair of \celerite\
terms with parameters
\begin{eqnarray}\eqlabel{sho-real}
a_{j\pm} &=& \frac{1}{2}\,S_0\,\omega_0\,Q\,\left[ 1 \pm
        \frac{1}{\sqrt{1-4\,Q^2}}\right] \\
b_{j\pm} &=& 0 \nonumber\\
    c_{j\pm} &=& \frac{\omega_0}{2\,Q}\,\left[1 \mp \sqrt{1-4\,Q^2}\right]
    \nonumber\\
d_{j\pm} &=& 0 \quad. \nonumber
\end{eqnarray}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/sho.pdf}
\caption{(left) The power spectrum of a stochastically-driven simple harmonic
    oscillator (\eqalt{sho-psd}) plotted for several values of the quality
    factor $Q$.
    For comparison, the dashed line shows the Lorentzian function from
    \eq{lorentz-psd} with $c_j = \omega_0/(2\,Q) = 1/20$ and normalized so that
    $S(d_j)/S(0) = 100$.
    (right) The corresponding autocorrelation functions with the same colors.
    \figurelabel{sho}}
\end{center}
\end{figure}

@@

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess\ \citep{Ricker:2014,Campante:2016}, we will infer
$\nu_\mathrm{max}$ and $\Delta\nu$ using only one month of \kepler\ data and
compare our results to the results inferred from the full 4 year baseline of
the \kepler\ mission.
For KIC~11615890, the published asteroseismic parameters based on several
years of \kepler\ observations are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
\citep{Dumusque:2014}.
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star \citep{Aigrain:2015} and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is everywhere positive.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.

%The general case with multiple \celerite\ terms can be checked using Sturm's
%theorem. The power spectrum is computed from taking the square of the Fourier
%transform of the data time series; hence, the power-spectrum must be non-negative.
%In constructing a kernel, if any of the parameters $a_j$ or $b_j$ are negative,
%then it is possible for the power spectrum to go negative, which violates the
%definition of the power spectrum.
%Consequently, if any of these coefficients is negative, it is necessary to
%check that the power spectrum is still non-negative.  Note that a non-negative
%power-spectrum does not require the auto-correlation
function to be positive for all $\tau$.  %The positive power spectrum is
%related to the positive eigenvalues of the covariance matrix which are
%necessary for a positive-definite matrix in limiting cases
%\citep{Messerschmitt:2006}.  %We find empirically that requiring an everywhere
%positive power spectrum results in positive eigenvalues for the covariance
%matrix, and so we describe here how to ensure a positive power spectrum using
%Sturm's theorem.
%Relation of positive power spectrum to positive definite covariance matrix.
%Power spectrum of a sum of kernels.

In the case of $J$ \celerite\ terms, we can check for negative values of the
PSD by solving for the roots of the power spectrum (see \eqalt{celerite-psd}),
abbreviating with $z = \omega^2$:
\begin{equation}
S(\omega)=  \sum_{j=1}^J \frac{q_j z + r_j}{z^2+s_jz + t_j} = 0
\end{equation}
where
\begin{eqnarray}
q_j &=& a_jc_j-b_jd_j\\
r_j &=& (d_j^2+c_j^2)(b_jd_j+a_jc_j)\\
s_j &=& 2(c_j^2-d_j^2)\\
t_j &=& (c_j^2+d_j^2)^2.
\end{eqnarray}
The denominators of each term are positive, so we can multiply through by
$\prod_j \left(z^2+s_jz + t_j\right)$ to find
\begin{equation}
Q_0(z) = \sum_{j=1}^J (q_j z + r_j)\prod_{k \ne j}\left(z^2+s_kz +
    t_k\right) = 0\quad,
\end{equation}
which is a polynomial with order $2\,(J-1)+1$.
With $J=2$, this yields a cubic equation which may be solved exactly for the
roots.

>>>>

To scale GP models to larger datasets, \citet{Rybicki:1995} presented a method
of computing the first term in \eq{gp-likelihood} in $\mathcal{O}(N)$
operations when the covariance function is given by
\begin{eqnarray}\eqlabel{kernel-simple}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} + a\,\exp(-c\,\tau_{nm})
\end{eqnarray}
where $\{{\sigma_n}^2\}_{n=1}^N$ are the measurement uncertainties,
$\delta_{nm}$ is the Kronecker delta, and $\bvec{\alpha} = (a,\,c)$.
The intuition behind this method is that, for this choice of $k_\bvec{\alpha}$,
the inverse of $K_\bvec{\alpha}$ is tridiagonal and can it can be computed
with a small number of operations for each data point.
Subsequently, \citet{Ambikasaran:2015} generalized this method to arbitrary
mixtures of exponentials
\begin{eqnarray}
k_\bvec{\alpha}(\tau_{nm}) = \sigma_n^2\,\delta_{nm} +
    \sum_{j=1}^J a_j\,\exp(-c_j\,\tau_{nm})\quad.
\end{eqnarray}
In this case, the inverse will be dense but \eq{gp-likelihood} can still be
evaluated in $\mathcal{O}(N)$ operations where $J$ is the number of components
in the mixture and $N$ is still the number of data points.

@@

In the case of the full \celerite\ covariance function
(\eqalt{celerite-kernel}), we introduce the following auxiliary variables in
analogy to the $u_n$ and $g_n$ that we introduced in the previous section
\begin{eqnarray}\eqlabel{full-system-first}
\phi_{n,j} &=& e^{-c_j\,\tau_{n+1,n}}\,\cos\left(d_j\,\tau_{n+1,n}\right)\\
\psi_{n,j} &=& -e^{-c_j\,\tau_{n+1,n}}\,\sin\left(d_j\,\tau_{n+1,n}\right)\\
g_{n,j} &=& \phi_{n,j}\,g_{n-1,j} + \phi_{n,j}\,z_n + \psi_{n,j}\,h_{n-1,j}\\
h_{n,j} &=& \phi_{n,j}\,h_{n-1,j} - \psi_{n,j}\,z_n - \psi_{n,j}\,g_{n-1,j}\\
u_{n,j} &=& \phi_{n+1,j}\,u_{n+1,j} + a_j\,z_{n+1} + \psi_{n+1,j}\,v_{n+1,j}\\
v_{n,j} &=& \phi_{n+1,j}\,v_{n+1,j} - b_j\,z_{n+1} - \psi_{n+1,j}\,u_{n+1,j}
\end{eqnarray}
with the boundary conditions
\begin{eqnarray}
    g_{0,j} = 0 \quad, \quad
    h_{0,j} = 0 \quad, \quad
    u_{N,j} = 0 \quad, \quad\mathrm{and}\quad
    v_{N,j} = 0
\end{eqnarray}
for all $j$.
Using these variables and some algebra, we find that the following expression
\begin{eqnarray}\eqlabel{full-system-last}
\sum_{j=1}^J \left[a_j\,g_{n,j}+b_j\,h_{n,j}\right]
+ \left[{\sigma_n}^2+\sum_{j=1}^J a_j\right]
+ \sum_{j=1}^J \left[\phi_{n,j}\,u_{n,j}+\psi_{n,j}\,v_{n,j}\right]
    &=& r_{\bvec{\theta},n}
\end{eqnarray}
is equivalent to the target matrix equation
\begin{eqnarray}
K\,\bvec{z} &=& \bvec{r}_\bvec{\theta}
\end{eqnarray}
if $r_{\bvec{\theta},n}$ is the $n$-th element of the residual vector
$\bvec{r}_\bvec{\theta}$ defined in \sect{gps}.
\eq{full-system-first} through \eq{full-system-last} define a banded matrix
equation in the ``extended'' space and, as before, this can be used to solve
for $K^{-1}\,\bvec{r}_\bvec{\theta}$ and $\det K$ in $\mathcal{O}(N)$
operations.
\Figure{matrix} shows a pictorial representation of the sparsity pattern of
the extended matrix $K_\mathrm{ext}$.
Given that definition of $K_\mathrm{ext}$, the corresponding extended vectors
$\bvec{z}_\mathrm{ext}$ and $\bvec{r}_\mathrm{ext}$ are defined schematically
as
\begin{eqnarray}
{\bvec{z}_\mathrm{ext}} ^\T =
\left(\begin{array}{cccccccccc}
    z_1 & u_{1,j} & v_{1,j} & g_{1,j} & h_{1,j} & z_2 & u_{2,j} & \cdots &
    h_{N-1,j} & z_N
\end{array}\right)
\end{eqnarray}
and
\begin{eqnarray}
{\bvec{r}_\mathrm{ext}} ^\T =
\left(\begin{array}{cccccccccc}
    r_{\bvec{\theta},1} & 0 & 0 & 0 & 0 & r_{\bvec{\theta},2} &
    0 & \cdots & 0 & r_{\bvec{\theta},2}
\end{array}\right) \quad.
\end{eqnarray}
After constructing the extended matrix (using a compact storage format), the
extended matrix can be factorized using a LU-decomposition\footnote{Even
though $K_\mathrm{ext}$ is symmetric, it is not positive definite so a
Cholesky solver cannot be used for increased efficiency.} routine optimized
for band or sparse matrices.
This decomposition can then be used to compute the determinant of $K$, solve
$K^{-1}\,\bvec{r}_\bvec{\theta}$, and subsequently calculate the marginalized
likelihood in \eq{gp-likelihood}.

@@

The extended system defined in the previous section is sparse with fewer than
a few percent non-zero entries and band structure.
In this \sectionname, we empirically investigate the performance and scaling
of three different algorithms for solving this extended system:

@@

\dfmtodo{Add more intro to this section.}

A special case of the \celerite\ model of great physical interest is a
stochastically-driven damped simple harmonic oscillator.
The differential equation for this system is
\begin{equation}
    \left[\frac{\dd^2}{\dd t^2} + \frac{\omega_0}{Q}\,\frac{\dd}{\dd t}
    + \omega_0^2\right]\, y(t) = \epsilon(t)
\end{equation}
where $\omega_0$ is the frequency of the undamped oscillator, $Q$ is the
quality factor of the oscillator, and $\epsilon(t)$ is a stochastic driving
force.
If $\epsilon(t)$ is white noise, the PSD of this process is given by
\citep[for example][]{Anderson:1990}
\begin{equation}\eqlabel{sho-psd}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0\,\omega_0^4}
    {(\omega^2-\omega_0^2)^2 + \omega_0^2\omega^2/Q^2}
\end{equation}
where $S_0$ is proportional to the power at $\omega = \omega_0$, $S(\omega_0)
= \sqrt{2/\pi}\,S_0\,Q^2$.
The power spectrum in \eq{sho-psd} matches \eq{celerite-psd} if
\begin{eqnarray}\eqlabel{sho-complex}
a_j &=& S_0\,\omega_0\,Q \\
b_j &=& \frac{S_0\,\omega_0\,Q}{\sqrt{4\,Q^2-1}} \\
c_j &=& \frac{\omega_0}{2\,Q}\\
d_j &=& \frac{\omega_0}{2\,Q} \sqrt{4\,Q^2-1} \quad,
\end{eqnarray}
for $Q \ge \frac{1}{2}$.
For $0 < Q \le \frac{1}{2}$, \eq{sho-psd} can be captured by a pair of \celerite\
terms with parameters
\begin{eqnarray}\eqlabel{sho-real}
a_{j\pm} &=& \frac{1}{2}\,S_0\,\omega_0\,Q\,\left[ 1 \pm
        \frac{1}{\sqrt{1-4\,Q^2}}\right] \\
b_{j\pm} &=& 0 \nonumber\\
    c_{j\pm} &=& \frac{\omega_0}{2\,Q}\,\left[1 \mp \sqrt{1-4\,Q^2}\right]
    \nonumber\\
d_{j\pm} &=& 0 \quad. \nonumber
\end{eqnarray}

@@

\begin{figure}[!htbp]
\begin{center}
\includegraphics[width=0.9\textwidth]{figures/sho.pdf}
\caption{(left) The power spectrum of a stochastically-driven simple harmonic
    oscillator (\eqalt{sho-psd}) plotted for several values of the quality
    factor $Q$.
    For comparison, the dashed line shows the Lorentzian function from
    \eq{lorentz-psd} with $c_j = \omega_0/2\,Q = 1/20$ and normalized so that
    $S(d_j)/S(0) = 100$.
    (right) The corresponding autocorrelation functions with the same colors.
    \figurelabel{sho}}
\end{center}
\end{figure}

@@

To demonstrate the applicability of this model, we apply it to infer the
asteroseismic parameters of the giant star KIC~11615890 observed by the
\kepler\ Mission.
The goal of this example is to show that, even for a low signal-to-noise
dataset with a short baseline, it is possible to infer asteroseismic
parameters with formal uncertainties that are consistent with the parameters
inferred with a much larger dataset.
Looking forward to \tess\ \citep{Campante:2016}, we will infer
$\nu_\mathrm{max}$ and $\Delta\nu$ using only one month of \kepler\ data and
compare our results to the results inferred from the full 4 year baseline of
the \kepler\ mission.
For KIC~11615890, the published asteroseismic parameters based on several
years of \kepler\ observations are \citep{Pinsonneault:2014}
\begin{eqnarray}
    \nu_\mathrm{max} = 171.94 \pm 3.62 \,\mu\mathrm{Hz} \quad\mathrm{and}\quad
    \Delta\nu = 13.28 \pm 0.29 \,\mu\mathrm{Hz} \quad.
\end{eqnarray}
We randomly select a month-long segment of \kepler\ data, initialize our
\celerite\ model using a grid search in the parameter space, and then use
\emcee\ \citep{Foreman-Mackey:2013} to sample the joint posterior density for
the full set of parameters.
\Figure{astero-corner} shows the marginalized density for $\nu_\mathrm{max}$
and $\Delta\nu$ compared to the results from the literature.
\dfmtodo{Add some more discussion here.}

@@

Another source of variability that can be measured from time series
measurements of stars is rotation.
The inhomogeneous surface of the star (spots, plage, \etc) imprints itself as
quasiperiodic variations in photometric or spectroscopic observations
(\dfmtodo{CITE}).
It has been demonstrated that for light curves with nearly uniform sampling,
the empirical autocorrelation function provides a reliable estimate of the
rotation period of a star (\dfmtodo{CITE}) and that a GP model with a
quasiperiodic covariance function can be used to make probabilistic
measurements even with sparsely sampled data (R.~Angus, \etal\ in prep.).
The covariance function used for this type of analysis has the form
\begin{eqnarray}\eqlabel{sine2}
k(\tau) = A\,\exp\left(-\frac{\tau^2}{2\,\ell^2} -
    \Gamma\,\sin^2\left(\frac{\pi\,\tau}{P} \right) \right)
\end{eqnarray}
where $P$ is the period of the oscillation.
The key difference between this function and other quasiperiodic kernels is
that it is everywhere positive.
We can construct a simple \celerite\ covariance function with similar properties
as follows
\begin{eqnarray}\eqlabel{rot-kernel}
k(\tau) = \frac{a}{2+b}\,e^{-c\,\tau}\,\left[
    \cos\left(\frac{2\,\pi\,\tau}{P}\right) + (1 + b)
\right]
\end{eqnarray}
for $a>0$, $b>0$, and $c>0$.
The covariance function in \eq{rot-kernel} cannot exactly reproduce \eq{sine2}
but, since \eq{sine2} is only an effective model, \eq{rot-kernel} can be used
as a drop-in replacement for a substantial gain in computational efficiency.

@@

Using \eq{celerite-psd}, we find that for a single \celerite\ term, this
requirement is met when
\begin{eqnarray}\eqlabel{psd-req}
\frac{(a_j\,c_j+b_j\,d_j)\,({c_j}^2+{d_j}^2)+(a_j\,c_j-b_j\,d_j)\,\omega^2}
{\omega^4+2\,({c_j}^2-{d_j}^2)\,\omega^2+{({c_j}^2+{d_j}^2)}^2} > 0 \quad.
\end{eqnarray}
The denominator is positive for all $c_j \ne 0$ and it can be shown that, when
$c_j=0$, \eq{psd-req} is satisfied for all $\omega \ne d_j$ where the power is
identically zero.
Therefore, when $c_j \ne 0$, we require that the numerator is positive for all
$\omega$.
This requirement can also be written as
\begin{eqnarray}
a_j\,c_j > -b_j\,d_j \\
a_j\,c_j > b_j\,d_j \quad.
\end{eqnarray}
Furthermore, we can see that $a_j$ must be positive since $k(0) = a_j$ should
should be positive and, similarly, by requiring the covariance to be finite at
infinite lag, we obtain the constraint $c_j \ge 0$.
Without loss of generality, we can choose $d_j \ge 0$ and find the final
constraint:
\begin{eqnarray}
    -|b_j\,d_j| < a_j\,c_j < |b_j\,d_j|
\end{eqnarray}

%The general case with multiple \celerite\ terms can be checked using Sturm's
%theorem. The power spectrum is computed from taking the square of the Fourier
%transform of the data time series; hence, the power-spectrum must be non-negative.
%In constructing a kernel, if any of the parameters $a_j$ or $b_j$ are negative,
%then it is possible for the power spectrum to go negative, which violates the
%definition of the power spectrum.
%Consequently, if any of these coefficients is negative, it is necessary to
%check that the power spectrum is still non-negative.  Note that a non-negative
%power-spectrum does not require the auto-correlation
%function to be positive for all $\tau$.  %The positive power spectrum is
%related to the positive eigenvalues of the covariance matrix which are
%necessary for a positive-definite matrix in limiting cases
%\citep{Messerschmitt:2006}.  %We find empirically that requiring an everywhere
%positive power spectrum results in positive eigenvalues for the covariance
%matrix, and so we describe here how to ensure a positive power spectrum using
%Sturm's theorem.
%Relation of positive power spectrum to positive definite covariance matrix.
%Power spectrum of a sum of kernels.

In the case of $J$ \celerite\ terms, we can check for negative values of the
PSD by solving for the roots of the power spectrum (see \eqalt{celerite-psd}),
abbreviating with $z = \omega^2$:
\begin{equation}
S(\omega)=  \sum_{j=1}^J \frac{q_j\,z + r_j}{z^2+s_j\,z + t_j} = 0
\end{equation}
where
\begin{eqnarray}
q_j &=& a_j\,c_j-b_j\,d_j\\
r_j &=& (d_j^2+c_j^2)(b_j\,d_j+a_j\,c_j)\\
s_j &=& 2(c_j^2-d_j^2)\\
t_j &=& (c_j^2+d_j^2)^2.
\end{eqnarray}
The denominators of each term are positive, so we can multiply through by
$\prod_j \left(z^2+s_jz + t_j\right)$ to find
\begin{equation}
Q_0(z) = \sum_{j=1}^J (q_j z + r_j)\prod_{k \ne j}\left(z^2+s_kz +
    t_k\right) = 0\quad,
\end{equation}
which is a polynomial with order $2\,(J-1)+1$.
With $J=2$, this yields a cubic equation which may be solved exactly for the
roots.
